# Data sources 

```{r}
library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)
library(devtools)
```

```{r}
data_by_day <- read_csv("data/data-by-day.csv")
head(data_by_day)
```

# first: the overall view of the summation over the 5 districts(BX, BK, MN, QN, SI)

```{r}
#data_transform <- data_by_day %>% pivot_longer(data_by_day, names_to = 'city', values_to = 'count')
overall_transform <- data_by_day  %>% pivot_longer(!date_of_interest, names_to = 'cases', values_to = 'count')

df <- overall_transform %>%
  mutate(areas = case_when(
    startsWith(cases, "BX_") ~ "BX",
    startsWith(cases, "BK_") ~ "BK",
    startsWith(cases, "MN_") ~ "MN",
    startsWith(cases, "QN_") ~ "QN",
    startsWith(cases, "SI_") ~ "SI",
    TRUE~ "sum_nyc"
    ))
df1 <- df %>% transform(cases=str_replace(cases,"BX_","")) %>% 
  transform(cases=str_replace(cases,"BK_","")) %>% 
  transform(cases=str_replace(cases,"MN_","")) %>%  
  transform(cases=str_replace(cases,"QN_","")) %>%  
  transform(cases=str_replace(cases,"SI_",""))
df1
```
#the final output data:

```{r}
df2 = df1 %>% pivot_wider(names_from = cases, values_from = count)
df2
```

# maybe wanna to drop the total_nyc value & include & compare only 5 districts

```{r}
df3 = df2 %>% filter(areas != "sum_nyc")
#drop the incomplete column
df3 <- df3[ -c(15) ]
```

# deal with the missing value/NA in the dataset & draw the plot to see the missing pattern:

```{r}
colSums(is.na(df3)) %>%
  sort(decreasing = TRUE)
```

#consider by the sum:
```{r}
library(dplyr)
total_sum <- select(data_by_day, 1:11)
total_sum_transform <- total_sum %>% pivot_longer(!date_of_interest, names_to = 'cases', values_to = 'count') %>% 
total_sum_transform
```
#By area:
```{r}
bx_sum <- select(data_by_day, 1, 12:22)
bx_transform <- bx_sum %>% 
  pivot_longer(!date_of_interest, names_to = 'cases', values_to = 'count') %>% 
  transform(cases=str_replace(cases,"BX_","")) 

bx_transform$Area <- "BX"
bx_transform
```

```{r}
#brooklyn area
bk_sum <- select(data_by_day, 1, 23:33)
bk_transform <- bk_sum %>% 
  pivot_longer(!date_of_interest, names_to = 'cases', values_to = 'count') %>% 
  transform(cases=str_replace(cases,"BK_",""))
bk_transform$Area <- "BK"
bk_transform
```

```{r}
#install.packages('data.table')
library(data.table)
setDT(bx_transform)
setDT(bk_transform)

bk_transform[bx_transform, on = c('date_of_interest','cases')]
```

```{r}
#manhatton
mn_sum <- select(data_by_day, 1, 34:44)
mn_transform <- mn_sum %>% 
  pivot_longer(!date_of_interest, names_to = 'cases', values_to = 'count') %>% 
  transform(cases=str_replace(cases,"MN_",""))
mn_transform$Area <- "MN"
mn_transform
```

```{r}
#queens
qn_sum <- select(data_by_day, 1, 45:55)
qn_transform <- qn_sum %>% 
  pivot_longer(!date_of_interest, names_to = 'cases', values_to = 'count') %>% 
  transform(cases=str_replace(cases,"QN_",""))
qn_transform$Area <- "QN"
qn_transform
```
```{r}
#si
si_sum <- select(data_by_day, 1, 56:66)
si_transform <- si_sum %>% 
  pivot_longer(!date_of_interest, names_to = 'cases', values_to = 'count') %>% 
  transform(cases=str_replace(cases,"SI_",""))
si_transform$Area <- "SI"
si_transform
```

# After modifying the dataset by area, we can join them together by date or analyze on each of it separately.

## the second dataset: now_weekly_breakthrough

```{r}
now_weekly_breakthrough <- read_csv("data/now-weekly-breakthrough.csv")
#head(now_weekly_breakthrough)
#install.packages('janitor')
library(janitor)

#can be done by simply
now_weekly_breakthrough <- clean_names(now_weekly_breakthrough)
now_weekly_breakthrough

```
```{r}
overall_transform1 <- now_weekly_breakthrough  %>% pivot_longer(!week_of_diagnosis, names_to = 'cases', values_to = 'count')
overall_transform1
```


```{r}
df5 <- overall_transform1 %>%
  mutate(vacci_status = case_when(
    startsWith(cases, "vaccinated_boosted") ~ "fully_vaccinated",
    startsWith(cases, "Vaccinated_unboosted") ~ "unboosted",
    startsWith(cases, "Unvaccinated") ~ "unvaccinated",
    TRUE~ "nyc_wide"
    ))

df6 <- df5 %>% transform(cases=str_replace(cases,"vaccinated_boosted_","")) %>% 
  transform(cases=str_replace(cases,"vaccinated_unboosted_","")) %>% 
  transform(cases=str_replace(cases,"Unvaccinated_","")) %>%  
  transform(cases=str_replace(cases,"citywide_",""))
df6
```

```{r}
df7 = df6 %>% filter(cases!="incomplete") 
df7
```
#there is no missing value:

```{r}
df7$count <- as.numeric(df7$count)
df7$count
```

```{r}
# by column
colSums(is.na(df7)) %>%
  sort(decreasing = TRUE)
```

