# Missing Value



```{r}
#install.packages("naniar")
library(naniar)
library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)
library(devtools)

```

All the missing values come from one column: *HOSP_COUNT_7DAY_AVG*. The total amount of missing value is 3935 for this dataset.

```{r}
data_by_day <- read_csv("data/data_by_day_transform_dataset.csv")
```

```{r, echo = TRUE}
colSums(is.na(data_by_day)) %>%
  sort(decreasing = TRUE)
```

Using groupby *areas* to see the distribution of missing value:

```{r, echo = T}
data_by_day %>% 
  select(HOSP_COUNT_7DAY_AVG, areas) %>%  
  group_by(areas) %>% miss_var_summary()
```



```{r, echo = T}
gg_miss_var(data_by_day,
            facet = areas)
```

The missing values are equally distributed among BX, BK, MN, QN, SI counties, which more likely to be happen on the same date.

```{r, echo = T}
missing_range <-data_by_day %>% 
  select(HOSP_COUNT_7DAY_AVG, date_of_interest) %>%  
  group_by(date_of_interest) %>% miss_var_summary()
```


```{r, echo = T}
head(missing_range)
```


```{r, echo = T}
tail(missing_range)
```

The missing value occurs from 02/29/2020 to 04/25/2022. Between this time frame, varaible *HOSP_COUNT_7DAY_AVG* records are missing over the 5 counties. 

To deal with those missing value, as we are not quite interested in variable *HOSP_COUNT_7DAY_AVG*, we will not be bothered by the existence of missing values and will keep all the records in the dataset to continue further analysis. 

Consider about the correlation between the variables in the dataset:

```{r, echo = T}
library(corrplot)
library(RColorBrewer)
source("http://www.sthda.com/upload/rquery_cormat.r")
mydata <- data_by_day[, c(3:8)]
require("corrplot")
rquery.cormat(mydata)
```
It is quite obvious to see that all cases are positive correlated and the correlation is quite strong for the count of cases and their 7 day count cases under the same category. 
